Prisoner's Dilemma

### 2020-04-23 realnotes2/notes.tex
modified theorems

\documentclass[11pt]{amsart}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{todonotes}

@@

\begin{definition}
  Suppose strategy $s_1$ plays strategy $s_2$. This defines a \textit{$s_1$-$s_2$ graph} which is a Markov chain where each node represents a pair of states $(c_1,c_2)$ where $c_1$ is a state in $s_1$ and $c_2$ is a state in $s_2$. The transition probabilities are defined in the obvious way.
\end{definition}

@@

Note: The graph might be periodic in which case we we will not get a stationary distribution. I need to think about this special case but my intuition is that it shouldn't matter.

@@

\begin{definition}
  A strategy $s_1$ is \textit{evolutionary stable} if there exists an $\alpha \in (0,1)$ such that for all $\epsilon < \alpha$, $s_1$ is not $\epsilon$-invadable.
\end{definition}

Note: it is easy to see that this is just equivalent to saying that there exists some $\epsilon$ for which $s_1$ is not $\epsilon$-invadable.

\begin{theorem}
  Suppose $s_1$ is evolutionary stable. Then $v_{s_1}(s_1) \geq \frac{S + T}{2}$.
\end{theorem}

\begin{theorem}
  Suppose $s_1$ is evolutionary stable as $p$ goes to 0 (i.e., that it is evolutionary stable if condition \ref{fitnesscond} is replaced by $\lim_{p \to 0} (F (s_2)-F(s_1)) > 0$). Then $v_{s_1}(s_1) = R$. In other words, $s_1$ is utilitarian.
\end{theorem}

\begin{theorem}
  The Pavlov strategy, displayed in Figure ??, is evolutionary stable as $p$ goes to 0.
\end{theorem}

\begin{remark}
  TFT, displayed in Figure ??, is not evolutionary stable as $p$ goes to 0. It has the stationary distribution $(1/4,1/4,1/4,1/4)$ which is smaller than $R$.
\end{remark}

\begin{figure}
  \includegraphics[width=4cm]{tft.jpg}
  \centering
  \caption{TFT.}
  \end{figure}
\begin{figure}
  \includegraphics[width=4cm]{pavlov.jpg}
  \centering
  \caption{Pavlov.}
  \end{figure}

>>>>

\documentclass[11pt]{amsart}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{todonotes}
\usepackage{cleveref}

@@

\begin{definition}
  Suppose strategy $s_1$ plays against strategy $s_2$. This defines an \textit{$s_1$-$s_2$ graph} which is a Markov chain where each node represents a pair of states $(c_1,c_2)$ where $c_1$ is a state in $s_1$ and $c_2$ is a state in $s_2$. The transition probabilities are defined in the obvious way.
\end{definition}

@@

Note: The graph might be periodic in which case we we will not get a stationary distribution — in that case, however, because of the linearity of the payoffs, we can replace the see that the periodic distribution gets the same time-average value as the stationary distribution that is the mean of the distributions of the periodic end state.

@@

\begin{definition}
  A strategy $s_1$ is \textit{evolutionarily stable} if there exists parameters $p_0$ and $\alpha$, both in $(0,1)$, such that for all $p < p_0$, and all $\epsilon < \alpha$, $s_1$ is not $\epsilon$-invadable.
\end{definition}

Note: it is easy to see that this is just equivalent to saying that for all $p < p_0$, there exists some $\epsilon$ for which $s_1$ is not $\epsilon$-invadable.

\begin{theorem}
  \label{evolutionarystable1}
  Suppose $s_1$ is evolutionarily stable. Then $\lim_{p \to 0} v_{s_1}(s_1) = R$.
\end{theorem}

\begin{theorem}
  \label{pavlovtheorem}
  The Pavlov strategy, displayed in \cref{pavlovfigure}, is evolutionarily stable.
\end{theorem}

\begin{figure}
  \label{tftfigure}
  \includegraphics[width=4cm]{tft.jpg}
  \centering
  \caption{TFT.}
\end{figure}
\begin{figure}
  \label{pavlovfigure}
  \includegraphics[width=4cm]{pavlov.jpg}
  \centering
  \caption{Pavlov.}
\end{figure}

\begin{remark}
  TFT, displayed in \cref{tftfigure}, is not evolutionarily stable. It has the stationary distribution $(1/4,1/4,1/4,1/4)$ which is smaller than $R$.
\end{remark}

    \begin{proof}[Proof of \cref{evolutionarystable1}]
      Suppose $s_1$ is such that it is not true that \begin{equation*}
        \lim_{p \to 0 } v_{s_1}(s_1) = R
      \end{equation*}
      Note first that $v_{s_1}(s_1)$ is a polynomial in $p$, so the limit exists \todo{Formally prove this!}. Also, by the symmetry of the game, we know that $v_{s_1}(s_1) \leq R$ is always true, so this must mean that $\lim_{p \to 0} v_{s_1}(s_1) < R$ \todo{Formally prove this!}. In particular, then, there exists some $\beta > 0$ such that $v_{s_1}(s_1) = R - \beta$.

      We want to prove that $s_1$ is not evolutionarily stable, and to do that, we want to prove that for any sufficiently small $\epsilon$, there exists a strategy $s_2$ that can invade $s_1$.

      We create the strategy $s_2$ as follows. First, copy the entire $s_1$ machine into $s_2$. Suppose the state corresponding to the start state of $s_1$ is $c_s$. Let the output at $c_s$ be $G(c_s)$. Let the node it goes to upon perceiving the opponent move $G(c_s)$ be $T(c_s, G(c_s))$. Then, create a new state $c_0$ that outputs $\lnot G(c_s)$ and has transition $T(c_0, G(c_s)) = T(c_s, G(c_s))$. Create another new state $c_1$. Let $T(c_0, \lnot G(c_s)) = c_1$. Let $T(c_2, \cdot) = c_2$. Let $G(c_2) = C$. This completely describes $s_2$.

      Now, we claim that the payoffs are as follows.
      \begin{equation}
        v_{s_2}(s_1) = (1 - p) \gamma  + p S
      \end{equation}
      \begin{equation}
        v_{s_2}(s_2) = (1-p)^{2} R + 2 (1-p) p (\tfrac{S + T}{2}) + p^2 \gamma
      \end{equation}
      \begin{equation}
        v_{s_1}(s_1) = \gamma
      \end{equation}
      \begin{equation}
        v_{s_1}(s_2) = (1-p) \gamma + p T
      \end{equation}

      Now, we simply compute $F(s_2) - F(s_1)$, and want to show that it is greater than 0 for all $\alpha$ that are sufficiently small.

      \begin{align*}
        F(s_2) - F(s_1) &= \\
        &= (1 - \alpha) \cdot v_{s_2}(s_1) + \alpha \cdot v_{s_2}(s_2) - (1 - \alpha) \cdot v_{s_1}(s_1) - \alpha \cdot v_{s_1}(s_2) \\
        &= 
      \end{align*}

      We can easily see that this works for some small $p$.

    \end{proof}

    \begin{proof}[Proof of \cref{pavlovtheorem}]

    \end{proof}
